---
title: "Statistical inference with the GSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup
In this project, the proportion of the men and women who have you been unemployed and looking for work for as long as a month.
The project was completed as a part of **Duke University’s ‘Inferential Statistics’** online course on **Coursera** platform.

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `gss`. Delete this note when before you submit 
your work. 


```{r load-data}
load("gss.Rdata")
```


* * *

## Part 1: Data
According to the data in the [General Social Survey](https://gss.norc.org/Get-Documentation) (GSS) is collected from an independently drawn sample of English-speaking persons (and Spanish-speaking persons, beginning from 2006) 18 years of age or above, living in non-institutional arrangements within the United States. Full probability sampling was employed in all surveys beginning from 1977.

Note that the data sample is generalizable as it collected randomly. And it was an observational 
study and not experimental so we cannot use it to determine causality between variables in the data. 

* * *

## Part 2: Research question
From the year 2000 to 2010, were men, of 18 years or above, more likely to have been unemployed and looking for work for as long as a month more than women?


The importance of unemployment is undeniable. Consequences may include homelessness,
crimes, families breakdown, etc. so if we know the chance of men and women for being unemployment,
the companies can arrange job positions more rationally related to employees' gender.

* * *

## Part 3: Exploratory data analysis
The GSS data will be filtered based on year which should be 2010 and ages equal
or greater than 18, also we will not select respondents with NA values for *unemp* field because they are not useful for purpose of this project.
  
* **unemp**: At any time during the last ten years, have respondent been unemployed and looking for work for as long as a month?
* **sex**: Respondent's sex

The data can be seen in the output of following code:
```{r data}
dataset <- gss %>%
  filter(!is.na(unemp), year==2010, age>=18) %>%
  select(sex, unemp)

str(dataset)
```
As we can see, *unemp* has a two levels including Yes and No values, in the next code snippet,
the figure indicate the data:
```{r}
ggplot(data = dataset, aes(x = unemp)) + 
  geom_bar() + ggtitle("Unemployment up to one month between 2000 and 2010") + 
  xlab("Unemploymnet status")
```
And we can see the situation of data according to both *unemp* and *sex*:
```{r}
dataset_tab <- table(dataset)
dataset_tab
```
It seems our female respondents are more than male ones in both group, either
unemployment up to one month in past 10 years or not.
But according to the following numbers the proportion of unemployed men is more than women.
```{r}
men_proportion <- 100 * dataset_tab[1] / (dataset_tab[1] + dataset_tab[3])
women_proportion <- 100 * dataset_tab[2] / (dataset_tab[2] + dataset_tab[4])
s <- c(men_proportion, women_proportion)
s
```
Unemployment proportion in this sample among the men is 38.72% while among the women is roughly 37.23%.

```{r}
dataset <- droplevels(dataset)

mosaicplot(prop.table(table(dataset),1), main = 'Sex vs Unemployment up to one month')
```

In the next part we will use statistical inference about proportion of male and female respondents
to find out the equality of unemployment in the whole population.

* * *

## Part 4: Inference

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.