---
title: "Statistical inference with the GSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup
In this project, the proportion of the men and women who have you been unemployed and looking for work for as long as a month.
The project was completed as a part of **Duke University’s ‘Inferential Statistics’** online course on **Coursera** platform.

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `gss`. Delete this note when before you submit 
your work. 


```{r load-data}
load("gss.Rdata")
```


* * *

## Part 1: Data
According to the data in the [General Social Survey](https://gss.norc.org/Get-Documentation) (GSS) is collected from an independently drawn sample of English-speaking persons (and Spanish-speaking persons, beginning from 2006) 18 years of age or above, living in non-institutional arrangements within the United States. Full probability sampling was employed in all surveys beginning from 1977.

Note that the data sample is generalizable as it collected randomly. And it was an observational 
study and not experimental so we cannot use it to determine causality between variables in the data. 

* * *

## Part 2: Research question
In the year 2010, were men, of 18 years or above, more likely to have been unemployed and looking for work for as long as a month more than women?

The importance of unemployment is undeniable. Consequences may include homelessness,
crimes, families breakdown, etc. so if we know the chance of men and women for being unemployment,
the companies can arrange job positions more rationally related to employees' gender.
* * *

## Part 3: Exploratory data analysis
  

* * *

## Part 4: Inference

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.